---
title: "Occupancy models for selected species"
author: "Giorgia Graells and Derek Corcoran"
date: "`r Sys.Date()`"
output:
  bookdown::github_document2:
    fig_caption: true
    toc: false
bibliography: Biblio.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Objective

This repository serves to document and store the analyses and results for the manuscript **Exploring habitat use of terrestrial and marine birds in urban coastal areas** sent to the jounrnal **Frontiers Ecology And Evolution**

## Methods

First we load the required packages

```{r LoadPackages, echo= T}
# For manipulating and reading raster datasets
library(raster)
library(terra)
# For cleaning datasets
library(tidyverse)
# For managing vector spatial datasets
library(sf)
# For caclulating occupancy models
library(unmarked)
# For selecting models
library(MuMIn)
```

Then we load the coordinates of the sampling sites for the surveys and transform them in to a `SpatVector` object:

```{r SamplingPoints}
Puntos_Hull <- read_csv("https://raw.github.com/derek-corcoran-barrios/LayerCreationBuffer/main/Coords.csv") %>% 
  mutate(geometry = str_remove_all(str_remove_all(str_remove_all(geometry, "c"), "\\("), "\\)")) 

Puntos_Hull$Lon <-  str_split(Puntos_Hull$geometry, pattern = ",", simplify = T)[,1] %>% as.numeric()
Puntos_Hull$Lat <-  str_split(Puntos_Hull$geometry, pattern = ", ", simplify = T)[,2] %>% as.numeric()

Puntos_Hull <- Puntos_Hull %>% 
  dplyr::select(-geometry) %>% 
  st_as_sf(coords = c(2,3), crs = 4326) %>% 
  st_transform(crs = "+proj=utm +zone=19 +south +datum=WGS84 +units=m +no_defs") %>%
  terra::vect()
```

Then we generate a vector of the distances used to calculate the proportion of landuse in meters as seen in @Graells_Genearation_of_layers_2022:

```{r Distances}
Distancias <- round(seq(from = 30, to = 5000, length.out = 10), -2)
Distancias[1] <- 30
```

We then download the rasters from that repository to generate the `Layers` list with one raster stack for each distance, and another list called `OccuVars` where we extract the values for the proportion of each landuse for each one of the points

```{r LayersFromDistances}
Layers <- list()
OccuVars <- list()

for(i in 1:length(Distancias)){
  Layers[[i]] <- terra::rast(paste0("/vsicurl/https://raw.github.com/derek-corcoran-barrios/LayerCreationBuffer/main/Proportions_", Distancias[i],".tif"))
  OccuVars[[i]] <- terra::extract(Layers[[i]], Puntos_Hull)
}
```

As an example in \@ref(fig:Layer600), we can see the proportions of each type of landuse for each point in the study site 600 meters around them.

```{r Layer600, fig.cap= "Here we see the proportion of each landuse 600 meters arround each point in the map", echo=FALSE}
plot(Layers[[2]])
```

Just as another example in Table \@ref(tab:Extract600) we see the extracted values for the first 10 sites of the study:

```{r Extract600, echo = FALSE}
knitr::kable(head(OccuVars[[i]], 10), caption = "The values of the proportion for the first ten sites of the study")
```

## Function for occupancy

We used the function `batchoccu2` which is a modification of the `batchoccu2`, from the DiversityOccupancy package [@Corcoran2021]. This function first fits all possible combinations of the probability of detection of a species, and selects the best model by AICc, and then using that model, for probability of detection, it tests all possible models for occupancy given that model and selects the best by AICc. In the next code 

```{r batchoccu2}
batchoccu2 <- function(pres, sitecov, obscov, spp, form, SppNames = NULL, dredge = FALSE) {
  if(is.null(SppNames)){
    SppNames <- paste("species", 1:spp, sep =".")
  }
  secuencia <- c(1:spp)*(ncol(pres)/spp)
  secuencia2<-secuencia-(secuencia[1]-1)
  models <- vector('list', spp)
  fit <- matrix(NA, nrow(pres), spp)
  Mods <- list()
  if(is.null(SppNames)){
    colnames(fit) <- paste("species", 1:spp, sep =".")
  }else if(class(SppNames) == "character"){
    colnames(fit) <- SppNames
  }
  if (dredge == FALSE) {
    for(i in 1:length(secuencia)) {
      data <- pres[, secuencia2[i]:secuencia[i]]
      data2 <- unmarkedFrameOccu(y = data, siteCovs = sitecov, obsCovs = obscov)
      try({
        models[[i]] <- occu(as.formula(form), data2)
      }, silent = T)
      try({
        fit[, i] <- suppressWarnings(predict(models[[i]], type = "state", newdata = sitecov))$Predicted
      }, silent = T)
      Mods = NULL
      print(paste("Species", as.character(i), "ready!"))
    }
  }
  else {
    for(i in 1:length(secuencia)) {
      data <- pres[, secuencia2[i]:secuencia[i]]
      data2 <- unmarkedFrameOccu(y = data, siteCovs = sitecov, obsCovs = obscov)
      try({
        #Partimos en dos Detección y occupancia
        form <- as.character(form)
        Div <- str_squish(form) %>% str_remove_all(" ")  %>% stringr::str_split(pattern = "~", simplify = T)
        
        ### Separamos dos formulas Occupancia y Deteccion
        
        Det <- Div[length(Div) - 1]
        
        VarDet <- str_split(Det, "\\+", simplify = T) %>% as.character()
        
        Fs <- list()
        
        print(paste("Starting to fit detection models for species", i, "of", length(secuencia)))
        
        for(x in 1:(length(VarDet) + 1)){
          if(x == (length(VarDet) + 1)){
            Formulas <- data.frame(Form = "~1 ~ 1", AICc = NA)
            Formulas$AICc[j] <- try(MuMIn::AICc(occu(as.formula("~1 ~1"), data2)), silent = T)
          }else{
            Test <- combn(VarDet, x, simplify = F)
            Formulas <- data.frame(Form = rep(NA, length(Test)), AICc = rep(NA, length(Test)))
            for(j in 1:length(Test)){
              Temp <- paste("~", paste(Test[[j]], collapse = " + "), "~ 1") 
              Formulas$Form[j] <- Temp
              Temp <- as.formula(Temp)
              Formulas$AICc[j] <- try(MuMIn::AICc(occu(Temp, data2)), silent = T) 
              gc()
            }
          }
          
          Fs[[x]] <- suppressWarnings(Formulas %>% mutate(AICc = as.numeric(AICc)) %>% dplyr::filter(!is.na(AICc)) %>% arrange(AICc))
          message(paste("finished for", x, "number of variables"))
        }
        
        Fs <- suppressWarnings(purrr::reduce(Fs, bind_rows) %>% arrange(AICc))
        
        Selected <- Fs$Form[1] %>% str_split("~", simplify = T) %>% as.character()
        Selected <- Selected[length(Selected) - 1] %>% str_squish()
        
        print(paste("Detection model for species", i, "is", Selected))
        
        Occup <- Div[length(Div)]
        
        VarOccup <- str_split(Occup, "\\+", simplify = T) %>% as.character()
        
        Fs <- list()
        
        print(paste("Starting to fit occupancy models for species", i, "of", length(secuencia)))
        
        for(x in 1:(length(VarOccup) + 1)){
          if(x == (length(VarOccup) + 1)){
            Formulas <- data.frame(Form = paste("~",Selected, "~ 1"), AICc = NA)
            Formulas$AICc[j] <- try(MuMIn::AICc(occu(as.formula(paste("~",Selected, "~ 1")), data2)), silent = T)
          }else{
            Test <- combn(VarOccup, x, simplify = F)
            Formulas <- data.frame(Form = rep(NA, length(Test)), AICc = rep(NA, length(Test)))
            for(j in 1:length(Test)){
              Temp <- paste("~", Selected, "~", paste(Test[[j]], collapse = " + ")) 
              Formulas$Form[j] <- Temp
              Temp <- as.formula(Temp)
              Formulas$AICc[j] <- try(MuMIn::AICc(occu(Temp, data2)), silent = T) 
              if((j %% 100) == 0){
                message(paste(j, "of", length(Test), "Ready"))
                gc()
              }
            }
          }
          
          Fs[[x]] <- suppressWarnings(Formulas %>% mutate(AICc = as.numeric(AICc)) %>% dplyr::filter(!is.na(AICc)) %>% arrange(AICc))
          message(paste("finished for", x, "number of variables", Sys.time()))
        }
        
        Fs <- suppressWarnings(purrr::reduce(Fs, bind_rows) %>% arrange(AICc))
        
        Mods[[i]] <- Fs
        
        
        Best <- Fs$Form[1]
        
        models[[i]] <- occu(as.formula(Best), data2)
        #dredged <- suppressWarnings(dredge(occu(form, data2)))
        # select the first model and evaluate
        #models[[i]] <- eval(getCall(dredged, 1))
        
      }, silent = T)
      try({
        #predictions for the best model
        fit[, i] <- suppressWarnings(predict(models[[i]], type = "state", newdata = sitecov))$Predicted
      }, silent = T)
      
      print(paste("Species", as.character(i), "ready!"))
    }
  }
  if(is.null(SppNames)){
    names(models) <- paste("species", 1:spp, sep =".")
  }else if(class(SppNames) == "character"){
    names(models) <- SppNames
  }
  
  if(is.null(SppNames)){
    names(Mods) <- paste("species", 1:spp, sep =".")
  }else if(class(SppNames) == "character" & !is.null(Mods)){
    names(Mods) <- SppNames
  }
  
  cond <- sapply(models, function(x) !is.null(x))
  models <- models[cond]
  fit <- fit[,cond]
  Not <- SppNames[!(cond)]
  if(sum(!cond) >= 1){
    message(paste("species", paste(Not, collapse = ", "), "did not converge, try with less variables"))
  }
  result <- list(Covs = sitecov, models = models, fit = fit, Mods = Mods)
  class(result)<- "batchoccupancy"
  return(result)
}
```


# References